---
title: "Solution - Predict how sales of weather-sensitive products are affected by snow and rain"
author: "ResearchBigD"
output: html_document
---

You are running this report at `r Sys.time()`


### Project details
#### Context
Walmart has provided daily unit sales of 111 items across 45 stores. These 45 stores are located at different places and are under 20 different weathers. A mapping of a store to the weather station is provided. Daily weather information is available for all the 20 weather stations.

#### Objective - Predict the amount of each product sold around the time of major weather events
1. **A weather event as any day in which more than an inch of rain or two inches of snow was observed**
2. **Predict the units sold for a window of ±3 days surrounding each storm**


###Significance of the solution to Retail industry
#### Retailers
Most of the large retailers have ~70% revenue locked in inventory. Any opportunity to predict sales and in-turn inventory could create a large impact on their financial performance.

Customer not being able to get required items to perpare for severe weather at their walmart may try other stores. This could lead to lower customer loyalty.

This solution will help retailers to predict better invenotry of items, ensure adequate inventory  of the items to sell at the time of severe weather and will leave them with lesser inventory to clear after the sever weather. 

#### Customers

#####***Code and technical details are not published in the report. You can see code at [github]()***

<!-- Load packages --> 
```{r echo=FALSE}
library(lubridate)
library(tidyr)
library(dplyr)
library(timeDate)
library(ggplot2)
```

<!-- Read training data --> 
```{r echo=FALSE}
Walmart.train <- read.csv("train.csv")
```

<!-- Initial structure of data --> 
```{r echo=FALSE, results='hide'}
str(Walmart.train)
```

<!-- 
data.frame':	4617600 obs. of  4 variables
$ date     : Factor w/ 1034 levels "2012-01-01"
$ store_nbr: int
$ item_nbr : int
$ units    : int
--> 

<!-- Convert date from factor to date in ymd format -->
```{r echo=FALSE, results='hide'}
Walmart.train$date <- ymd(Walmart.train$date)
```

<!-- Structure of data after conversion to date --> 
```{r echo=FALSE, results='hide'}
str(Walmart.train)
```

<!-- 
data.frame':	4617600 obs. of  4 variables
$ date     : Date, format: "2012-01-01"
$ store_nbr: int
$ item_nbr : int
$ units    : int
--> 

<!-- Structure of data after conversion to date --> 
```{r echo=FALSE, results='hide'}
summary(Walmart.train$date)
```

<!-- 
min - 2012-01-01
Max - 2014-10-31
Days = 1034
--> 

<!-- 
1 - Add column - Is MonToThu
2 - Add column - Is FriToSun
3 - Add column - Is Holiday
--> 

```{r echo=FALSE}
Walmart.train$IsMonToThu <- as.integer(isWeekday(Walmart.train$date, wday = 1:4))
Walmart.train$IsFriToSun <- as.integer(isWeekend(Walmart.train$date, wday = 1:4))
```

<!-- Following code takes 12 hours to run --> 
```{r echo=FALSE, eval= FALSE}
Walmart.train$HolidayShopping <- isHolidayShoppingDay(Walmart.train$date)
Walmart.train <- read.csv("Walmart_HolidayShopping.csv")
Walmart.train$date <- ymd(Walmart.train$date)
```

```{r echo= FALSE}
Walmart.train.GroupByStore <- Walmart.train  %>% group_by(item_nbr, store_nbr)  %>% summarise(unit_sold = sum(units))

ggplot(subset(Walmart.train.GroupByStore, item_nbr == 1), aes(x = store_nbr, y = unit_sold)) + geom_line() + ggtitle("Units sold of the item by each store")
ggplot(subset(Walmart.train.GroupByStore, item_nbr == 2), aes(x = store_nbr, y = unit_sold)) + geom_line() + ggtitle("Units sold of the item by each store")

```

```{r echo=FALSE}
ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 0 & item_nbr <= 10 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 10 & item_nbr <= 20 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 20 & item_nbr <= 30 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 30 & item_nbr <= 40 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 40 & item_nbr <= 50 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 50 & item_nbr <= 60 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 60 & item_nbr <= 70 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 70 & item_nbr <= 80 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 80 & item_nbr <= 90 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 90 & item_nbr <= 100 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

ggplot(subset(Walmart.train.GroupByStore, (item_nbr > 100 & item_nbr <= 111 & unit_sold > 0)), aes(x = store_nbr, y = factor(item_nbr), size = unit_sold)) + geom_point(shape = 1) + ggtitle("Units sold of the item by each store")

countOfStoreSellingTheItem <- subset(Walmart.train.GroupByStore, unit_sold > 0)  %>% group_by(item_nbr)  %>% summarise(sellingStore = n())  %>% arrange(sellingStore)

ggplot(countOfStoreSellingTheItem, aes(x = factor(sellingStore), y = factor(item_nbr))) + geom_jitter(shape = 1, size = 2)

countOfItemsSoldInStore <- subset(Walmart.train.GroupByStore, unit_sold > 0)  %>% group_by(store_nbr)  %>% summarise(uniqItemsSold = n())  %>% arrange(uniqItemsSold)

ggplot(countOfItemsSoldInStore, aes(x = factor(uniqItemsSold), y = factor(store_nbr))) + geom_point(shape = 1, size = 2)




```

<!-- review items being sold in more than 10 stores -->
```{r echo=FALSE}
countOfStoreSellingTheItem  %>% arrange(desc(sellingStore))
item93 <- subset(Walmart.train, item_nbr == 93 & units > 0)
item5 <- subset(Walmart.train, item_nbr == 5 & units > 0)
item9 <- subset(Walmart.train, item_nbr == 9 & units > 0)
item45 <- subset(Walmart.train, item_nbr == 45 & units > 0)

ggplot(item93, aes(y = units, x = date, color = factor(store_nbr), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item93")
ggplot(item5, aes(y = units, x = date, color = factor(store_nbr), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5)+ ggtitle("Item5")
ggplot(item9, aes(y = units, x = date, color = factor(store_nbr), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5)+ ggtitle("Item9")
ggplot(item45, aes(y = units, x = date, color = factor(store_nbr), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5)+ ggtitle("Item45")
```

<!-- review items being sold in more than 10 stores for weekday and weekend -->
```{r echo=FALSE}
ggplot(item93, aes(y = units, x = date, color = factor(IsMonToThu), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item93")
ggplot(item5, aes(y = units, x = date, color = factor(IsMonToThu), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item5")
ggplot(item9, aes(y = units, x = date, color = factor(IsMonToThu), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item9")
ggplot(item45, aes(y = units, x = date, color = factor(IsMonToThu), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item45")

```

<!-- review items being sold for holiday and non holiday -->
```{r echo=FALSE}
ggplot(item93, aes(y = units, x = date, color = factor(HolidayShopping), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item93")

ggplot(item5, aes(y = units, x = date, color = factor(HolidayShopping), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item5")

ggplot(item9, aes(y = units, x = date, color = factor(HolidayShopping), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item9")

ggplot(item45, aes(y = units, x = date, color = factor(HolidayShopping), label = store_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("Item45")


```

<!-- Review stores -->
```{r echo=FALSE}
countOfItemsSoldInStore  %>% arrange(desc(uniqItemsSold))
store1 <- subset(Walmart.train, store_nbr == 1 & units > 0)
store1_full <- subset(Walmart.train, store_nbr == 1)

store22 <- subset(Walmart.train, store_nbr == 22 & units > 0)
store37 <- subset(Walmart.train, store_nbr == 37 & units > 0)
store31 <- subset(Walmart.train, store_nbr == 31 & units > 0)
```

<!-- Review sales pattern, if all items have average pattern -->
```{r echo=FALSE}
ggplot(store1, aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store1")

ggplot(store1_full, aes(y = units, x = date, color = factor(item_nbr))) + geom_line() + ggtitle("store1_full")

ggplot(store22, aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store22")

ggplot(subset(store22, item_nbr == 93), aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store22")

ggplot(store37, aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store37")

ggplot(subset(store37, item_nbr == 5), aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store37")

ggplot(store31, aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store31")

ggplot(subset(store31, item_nbr == 45), aes(y = units, x = date, color = factor(item_nbr), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5) + ggtitle("store31")

```

<!-- Review sales pattern, if customers buy more on weekday or weekend or holiday -->
```{r echo=FALSE}
ggplot(store1, aes(y = units, x = date, color = factor(IsMonToThu), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5)  + ggtitle("store1")

ggplot(store22, aes(y = units, x = date, color = factor(IsMonToThu), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5)  + ggtitle("store22")
ggplot(store37, aes(y = units, x = date, color = factor(IsMonToThu), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5)  + ggtitle("store37")
ggplot(store31, aes(y = units, x = date, color = factor(IsMonToThu), label = item_nbr)) + geom_line() + geom_text(nudge_x = 5)  + ggtitle("store31")

ggplot(store1, aes(y = units, x = date, color = factor(HolidayShopping), label = item_nbr)) + geom_text(nudge_x = 5)  + geom_line() + ggtitle("store1")

ggplot(store22, aes(y = units, x = date, color = factor(HolidayShopping), label = item_nbr)) + geom_text(nudge_x = 5)  + geom_line() + ggtitle("store22")

ggplot(store37, aes(y = units, x = date, color = factor(HolidayShopping), label = item_nbr)) + geom_text(nudge_x = 5)  + geom_line() + ggtitle("store37")

ggplot(store31, aes(y = units, x = date, color = factor(HolidayShopping), label = item_nbr)) + geom_text(nudge_x = 5)  + geom_line() + ggtitle("store31")

```

<!-- Review items with highest quantity -->
```{r echo=FALSE}
highestSellingItems <- subset(Walmart.train, units > 0) %>% group_by(item_nbr) %>% summarise(total_units = sum(units)) %>% arrange(desc(total_units))
head(highestSellingItems)
ggplot(highestSellingItems, aes(y = total_units, x = item_nbr, size = total_units, label = item_nbr)) + geom_point() + geom_text(nudge_x = 5)

ggplot(item45, aes(y = units, x = factor(store_nbr))) + geom_boxplot() + ggtitle("item45")

ggplot(item9, aes(y = units, x = factor(store_nbr))) + geom_boxplot() + ggtitle("item9")

ggplot(item5, aes(y = units, x = factor(store_nbr))) + geom_boxplot() + ggtitle("item5")

item44 <- subset(Walmart.train, item_nbr == 44 & units > 0)
ggplot(item44, aes(y = units, x = factor(store_nbr))) + geom_boxplot() + ggtitle("item44")




```

<!-- review all items -->
```{r echo=FALSE}
ggplot(subset(Walmart.train, item_nbr == 5), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item5")
ggplot(subset(Walmart.train, item_nbr == 6), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item6")
ggplot(subset(Walmart.train, item_nbr == 8), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item8")
ggplot(subset(Walmart.train, item_nbr == 9), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item9")
ggplot(subset(Walmart.train, item_nbr == 15), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item15")
ggplot(subset(Walmart.train, item_nbr == 16), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item16")
ggplot(subset(Walmart.train, item_nbr == 23), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item23")
ggplot(subset(Walmart.train, item_nbr == 25), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item25")
ggplot(subset(Walmart.train, item_nbr == 28), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item28")

ggplot(subset(Walmart.train, item_nbr == 44), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item44")

ggplot(subset(Walmart.train, item_nbr == 45), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item45")

ggplot(subset(Walmart.train, item_nbr == 93), aes(y = units, x = factor(store_nbr), label = units)) + geom_boxplot() + geom_text(nudge_x = .5) + ggtitle("item93")

```
[Kaggle link for project details](https://www.kaggle.com/c/walmart-recruiting-sales-in-stormy-weather)



