---
title: "My approach for Data Wrangling Exercise 1- Basic Data Manipulation"
author: "ResearchBigD"
date: "December 7, 2016"
output: html_document
---

You are running this report at `r Sys.time()`

#How to run this report and get final clean data in one step? I have to run it step by step to get the final output#

#1: Set-up

**I run this code to start execution on clean environment**

A Remove all objects in worspace

rm(list = ls())

B Set working directory

setwd("C:/Users/Prabhat_nonadmin/Desktop/BigData/R_Programming/WorkingDirectory")

**Load required packages at start**

C Load packages

```r
library(readr)
library(dplyr)
library(gdata)
```

#2: Load the data in RStudio

```r
refine_orig <- read_csv("refine_original.csv")
```

```r
glimpse(refine_orig)
```

#3: Clean up brand names
Clean up the 'company' column, so all of the misspellings of the brand names are standardized. For example, you can transform the values in the column to be: philips, akzo, van houten and unilever (all lowercase).

###3.1 - Convert into lower case
```r
refine_comp_lower <- refine_orig %>% mutate(comp_lower = tolower(company))
```
###3.2 - Do spell correction
```r
refine_philips_vector <- gsub("^[pf].*ps$","philips",refine_comp_lower$comp_lower)
refine_akzo_vector <- gsub("^ak.*$","akzo",refine_philips_vector)
refine_unilever_vector <- gsub("^un.*$","unilever",refine_akzo_vector)
refine_comp_lower2 <- refine_comp_lower %>% mutate(comp_lower2 = refine_unilever_vector)

```

#4: Separate product code and number
Separate the product code and product number into separate columns i.e. add two new columns called product_code and product_number, containing the product code and number respectively

```r
#Use seperate function of tidyr
library(tidyr)
refine_sep_prod_num <- refine_comp_lower2 %>% separate(`Product code / number`,c("product_code","product_number"),'-')
```



```{r echo = FALSE, eval=FALSE}
#Below is for my reference. Code will not show up or run
refine_sep_prod_num <- refine_comp_lower2 %>% mutate(product_code = gsub("([^.*]+)-.*$","\\1",`Product code / number`), product_number = gsub("^.*-","",`Product code / number`))
gsub("([^.*]+)-.*$","\\1",temp_string)
gsub("^.*-","",temp_string)
[1] "def"
gsub("-.*$","",temp_string)
[1] "abc"
```

#5: Add product categories
You learn that the product codes actually represent the following product categories:

* p = Smartphone
* v = TV
* x = Laptop
* q = Tablet

In order to make the data more readable, add a column with the product category for each record.

###Option # 1
```
category_xref <- data.frame(key = c("p","v","x","q"), val = c("Smartphone", "TV", "Laptop", "Tablet"))

library(qdapTools)

refine_prod_cat <- refine_sep_prod_num %>% mutate(product_category = lookup(product_code,category_xref[1:4,]))
```

###Option # 2 - *Code is for review/reference, but will not run*

```{r eval=FALSE}
prod_code_factor <- factor(refine_sep_prod_num$product_code)
levels(prod_code_factor) <- c("Smartphone", "Tablet", "TV", "Laptop")
refine_prod_cat <- refine_sep_prod_num %>% mutate(product_category = prod_code_factor)

```

###Option # 3 - *Code is for review/reference, but will not run*

```{r eval=FALSE}
tut <- c("p" = "Smartphone", "v" = "Tablet", "x" = "TV", "q" = "Laptop")
refine_prod_cat <- refine_sep_prod_num %>% mutate(product_category = tut[product_code])
```

#6: Add full address for geocoding
You'd like to view the customer information on a map. In order to do that, the addresses need to be in a form that can be easily geocoded. Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

```r
refine_full_addr <- refine_prod_cat %>% mutate(full_address = paste(address, city, country, sep = ","))
```

#7: Create dummy variables for company and product category
Both the company name and product category are categorical variables i.e. they take only a fixed set of values. In order to use them in further analysis you need to create dummy variables. Create dummy binary variables for each of them with the prefix company_ and product_ i.e.

###7.1 - Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever

```r
refine_bin_col <- refine_full_addr %>% mutate(company_philips = ifelse(comp_lower2 == "philips",1,0)) %>% mutate(company_akzo = ifelse(comp_lower2 == "akzo",1,0)) %>% mutate(company_van_houten = ifelse(comp_lower2 == "van houten",1,0)) %>% mutate(company_unilever = ifelse(comp_lower2 == "unilever",1,0))
```
###7.2 - Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet

```r
refine_bin_col2 <- refine_bin_col %>% mutate(product_smartphone = ifelse(product_category == "Smartphone",1,0)) %>% mutate(product_tv = ifelse(product_category == "TV",1,0)) %>% mutate(product_laptop = ifelse(product_category == "Laptop",1,0)) %>% mutate(product_tablet = ifelse(product_category == "Tablet",1,0))
```

#8: Submit the project on Github
Include your code, the original data as a CSV file refine_original.csv, and the cleaned up data as a CSV file called refine_clean.csv.


```r
refine_clean <- refine_bin_col2 %>% select(comp_lower2, product_code, product_number, full_address, name,company_philips, company_akzo, company_van_houten, company_unilever, product_smartphone, product_tv, product_laptop, product_tablet )
```

```r
glimpse(refine_clean)
```

```r
write.csv(refine_clean, file = "refine_clean.csv")
```

# Please review output refine_clean.csv in your working directory